---
title: "Applying Statistical Analysis"
subtitle: Prediction of restaurant inspections in NYC
output:
  pdf_document:
    number_sections: true
  html_document:
    code_folding: hide
    highlight: espresso
    number_sections: yes
    theme: journal
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(gridExtra)
# Reading in tame data
df_train <- read_rds("df_train.rds")
df_features <- read_rds("df_features.rds")
lin_reg_model <- read_rds("lin_reg_model.rds")
```

# Introduction

In the following, I will describe two instances where I used statistical analysis in my project. First, there is an exploration between one of the predictors, namely *grade* and the target, *days_until_next_inspection*. Next, I will carry out a similar descriptive analysis for *violation_group* and the target. 

# Days vs. grade

```{r}
df_train %>%
  ggplot ( aes ( x = grade, y = as.numeric ( days_until_next ) ) ) +
  geom_boxplot ( fill = "grey") + 
  geom_jitter(alpha = .1, width = .07) +
  xlab("grade") +
  ylab("days until next inspection") +
  theme(
    text = element_text(
      family = "serif",
      color = "gray25"
    )
  )
```

Here, we see boxplots of the target feature for each level of *grade*. I added some random 'horizontal' noise to the observations as a remedy against overplotting.The median number of days decreases as we move from A to C.

# Days vs. violation groups

```{r}
# Plotting function
plot_vs_days <- function(z) {
  
  df_train %>% 
    ggplot(aes(x = as.factor(z), y = as.numeric(days_until_next))) + 
    geom_boxplot() +
    xlab("Number of violations") +
    ylab("") +
    theme(
      text = element_text(
        family = "serif",
        color = "gray25"
        )
      )
  
}

# Plot objects
plot_lst <- df_train %>% 
  select(starts_with("viol_")) %>% 
  map(plot_vs_days)

# Arrange plots in grid
grid.arrange(
  arrangeGrob(plot_lst[[1]], top = "Vermin") , 
  arrangeGrob(plot_lst[[2]], top = "Facility"),
  arrangeGrob(plot_lst[[3]], top = "Food"),
  arrangeGrob(plot_lst[[4]], top = "Hygiene"), 
  arrangeGrob(plot_lst[[5]], top = "Not scored"), 
  ncol = 2)

# Remove auxiliary objects
rm(plot_lst)
rm(plot_vs_days)
```

The median number of days until the next inspection decreases as the number of violations per inspection increases.